---
title: "Asthma data set exploration"
output: html_notebook
---

## Discovery of the data

```{r}
wd <- "/data/Datasets_Local/Asthma"
```
I used the shell command: >head --lines 2 mrna.tsv to have a primary view of the data

There is headers and it is tab separated

Let's explore the data:

```{r}
path1 <-  paste0(wd,"/pagapow-DataExport-119281/subset1_UBIO_ADULT_FINAL/mrna.tsv")
gene.data1 <- read.delim(file = path1, header = TRUE, sep = "\t")
```

```{r}
head(gene.data1)
```

```{r}
clinical_data <- read.delim(paste0(wd,"/pagapow-DataExport-119385/subset1_UBIO_ADULT_FINAL/Clinical/clinical_i2b2trans.txt"))
```

##mRNA-like data

We want to understand how is organised the data. For the moment, it's like the result of a sql joint. We want to have a table for each datatype.

We'll start to work with genomic data (gene.data1) with a reduced set of patient.

```{r}
permut <- order(as.character(gene.data1$PATIENT.ID))
sorted.gene.data1 <- gene.data1[permut,]
#We add a column for the reduction:
facToNum <- function(f) {
  return(as.integer(substring(as.character(f), 3, 5))) #A_023 becomes 23
}
sorted.gene.data1$int.patientID <- unlist(lapply(sorted.gene.data1$PATIENT.ID, 
                                      facToNum))

```

```{r}
selection <- function(sorted.data, patientList) {
  #numList contains a list of integers 
  return(sorted.data[sorted.data$int.patientID %in% patientList, ])
}
```

```{r}
onePatient <- selection(sorted.gene.data1, 0)
dim(onePatient)[1]
```

Have they all the same number of features ?

```{r}
patientsNumbers <- unique(sorted.gene.data1$int.patientID) #list of patients id in integer format
listNumberFeatures <- c()
for (i in patientsNumbers) {
  listNumberFeatures <- c(listNumberFeatures, dim(selection(sorted.gene.data1, i))[1]) #number of features
}
length(unique(listNumberFeatures))
```

Let's make a matrix with one patient per row and the features in columns. 
We want one matrix by datatype. WE'LL CONSIDER THAT THE DATATYPE IS THE TISSUE TYPE HERE.
So here, it's only one data type.

Let's create a function to split data by a specified row (TISSUE.TYPE if our assumption is good)


```{r}
split_by_type <- function(data,data_column) {
  listOfType <- unique(data_column)
  resList <- list()
  for(type in listOfType) {
    resList <- append(resList, data[data_column == type,])
  }
  return(resList)
}

matrixMaker <- function(oneDataType) { #the dataframe is supposed to have the added column int.patientID
  listPatientID <- unique(oneDataType$int.patientID)
  firstIteration <- TRUE
  for (i in listPatientID) {
    #data from patient i in temp
    temp <- selection(oneDataType, i)
    one_row <- t(as.matrix(temp$VALUE))
    
    #columns named by the GENE.ID
    cnames <- t(temp$GENE.ID)
    colnames(one_row) <- cnames
    
    #patient identifier in integer form (can be change later if not convenient)
    rownames(one_row) <- i 
    
    if (firstIteration) {
      print(dim(one_row))
      resMatrix <- one_row #for the first iteration
      print(dim(resMatrix))
      firstIteration <- FALSE
    } else {
      resMatrix <- rbind(resMatrix, one_row)  # add the row to the whole matrix
    }
  }
  return(resMatrix)# a matrix with all the values of the data frame for each patient.
}


matrix1 <- matrixMaker(sorted.gene.data1)
```

```{r}
mostFeaturedPatient <- function(oneDataType) {
  patientsNumbers <- unique(oneDataType$int.patientID)
  numbMax <- 0
  for (i in patientsNumbers) {
    current_number <- dim(selection(oneDataType, i))[1] #number of features
    if (current_number > numbMax) {numbMax <- current_number}
}
  return(numbMax)
}
```

```{r}
matrix_creator <- function(oneDataType) {
  #sort the data by patients and add a column with patient ids as integer:
  permut <- order(as.character(oneDataType$PATIENT.ID))
  sorted.data <- oneDataType[permut,]
  facToNum <- function(f) {
    return(as.integer(substring(as.character(f), 3, 5))) #A_023 becomes 23
  }
  sorted.data$int.patientID <- unlist(lapply(sorted.data$PATIENT.ID, 
                                        facToNum))
  
  #init the matrix with the patient with the more features.
  most <- mostFeaturedPatient(sorted.data) 
  temp <- selection(sorted.data, most)
  col <- temp$VALUE
  print(dim(col))
  colnames(col) <- sorted.data$PATIENT.ID[most]
  rownames(col) <- as.vector(sorted.data$GENE.ID) 
  resMatrix <- col
  print(head.matrix(resMatrix))
}
matrix_creator(gene.data1)
```

```{r}
#create the list of all features (all different GENE.ID)
patientsNumbers <- unique(sorted.gene.data1$int.patientID) #list of patients id in integer format
allFeatures <- list()
for (i in patientsNumbers) {
  current_list <- selection(sorted.gene.data1, i)
  
}
```


